---
title: "Getting started with vroom"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{get-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The vroom package contains one function `vroom()` which is used to
read all types of delimited files. A delimited file is any file in
which the data is separated (delimited) by one or more characters.

The most common type of delimited files are CSV (Comma Separated
Values) files, typically these files have a `.csv` suffix.

## reading CSV files
To read a CSV file with vroom specify `delim = ","` to the call to vroom.

We have included an example CSV file in the vroom package. Access it with
`vroom_example("mtcars.csv")`

```{r setup}
library(vroom)

# See where the example file is stored on your machine
file <- vroom_example("mtcars.csv")
file

# Read the file, by default vroom will guess the delimiter automatically.
vroom(file)

# You can also specify it explicitly, which is (slightly) faster, and safer if
# you know how the file is delimited.
vroom(file, delim = ",")
```
