---
title: "Get started with vroom"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{get-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = tempdir())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(tibble.print_min = 3)
```

The vroom package contains one main function `vroom()` which is used to
read all types of delimited files. A delimited file is any file in
which the data is separated (delimited) by one or more characters.

The most common type of delimited files are CSV (Comma Separated
Values) files, typically these files have a `.csv` suffix.

```{r}
library(vroom)
```

This vignette covers the following topics:

- The basics of reading files, including
  - single files
  - multiple files
  - compressed files
  - remote files
- Skipping particular columns.
- Specifying column types, for additional safety and when the automatic guessing
  fails.

## Reading files
To read a CSV, or other type of delimited file with vroom pass the file to
`vroom()`. The delimiter will be automatically guessed if it is a common
delimiter. If the guessing fails or you are using a less common delimiter
specify it with the `delim` parameter. (e.g. `delim = ","`).

We have included an example CSV file in the vroom package for use in examples
and tests. Access it with `vroom_example("mtcars.csv")`

```{r}
# See where the example file is stored on your machine
file <- vroom_example("mtcars.csv")
file

# Read the file, by default vroom will guess the delimiter automatically.
vroom(file)

# You can also specify it explicitly, which is (slightly) faster, and safer if
# you know how the file is delimited.
vroom(file, delim = ",")
```

## Reading multiple files

If you are reading a set of files which all have the same columns, you can pass
the filenames directly to `vroom()` and it will combine them into one
result.

First we will create some files to read by splitting the mtcars dataset by
number of cylinders, (it is OK if you don't currently understand this code).
```{r}
mt <- tibble::rownames_to_column(mtcars, "model")
purrr::iwalk(
  split(mt, mt$cyl),
  ~ vroom_write(.x, glue::glue("mtcars_{.y}.csv"), "\t")
)
```

We can then efficiently read them into one table by passing the filenames directly
to vroom.

```{r}
files <- fs::dir_ls(glob = "mtcars*csv")
files
vroom(files)
```

Often the filename or directory where the files are stored contains
information, in this case the `id` parameter can be used to add an extra column
to the result with the full path to each file. (in this case named `path`).

```{r}
vroom(files, id = "path")
```

```{r, include = FALSE}
gc()
fs::file_delete(files)
```

## Reading compressed files

vroom supports reading zip, gz, bz2 and xz compressed files automatically,
just pass the filename of the compressed file to vroom.

```{r}
file <- vroom_example("mtcars.csv.gz")

vroom(file)
```

## Reading remote files

vroom can read files from the internet as well by passing the URL of the file
to vroom.

```{r, eval = as.logical(Sys.getenv("NOT_CRAN", "false"))}
file <- "https://raw.githubusercontent.com/r-lib/vroom/master/inst/extdata/mtcars.csv"
vroom(file)
```

It can even read gzipped files from the internet (although currently not the other
compressed formats).

```{r, eval = as.logical(Sys.getenv("NOT_CRAN", "false"))}
file <- "https://raw.githubusercontent.com/r-lib/vroom/master/inst/extdata/mtcars.csv.gz"
vroom(file)
```

## Column selection

vroom provides the same interface for column selection and renaming as
[dplyr::select()]. This provides very flexible and readable selections.

- A character vector of column names
```{r}
file <- vroom_example("mtcars.csv.gz")

vroom(file, col_select = c(model, cyl, gear))
```
- A numeric vector of column indexes, e.g. `c(1, 2, 5)`
```{r}
vroom(file, col_select = c(1, 3, 11))
```
- You can also use the selection helpers
```{r}
vroom(file, col_select = starts_with("d"))
```
- Or rename columns
```{r}
vroom(file, col_select = list(car = model, everything()))
```

## Reading fixed width files

 A fixed width file can be a very compact representation of numeric
 data. Unfortunately, it's also painful because you need to describe the length
 of every field. vroom aims to make it as easy as possible by providing a
 number of different ways to describe the field structure. Use `vroom_fwf()` in
 conjunction with one of the following helper functions to read the file.

```{r}
fwf_sample <- vroom_example("fwf-sample.txt")
cat(readLines(fwf_sample))
```

- `fwf_empty()` - Guess based on the position of empty columns.
```{r}
vroom_fwf(fwf_sample, fwf_empty(fwf_sample, col_names = c("first", "last", "state", "ssn")))
```

- `fwf_widths()` - Use user provided set of field widths.
```{r}
vroom_fwf(fwf_sample, fwf_widths(c(20, 10, 12), c("name", "state", "ssn")))
```

- `fwf_positions()` - Use user provided sets of start and end positions.
```{r}
vroom_fwf(fwf_sample, fwf_positions(c(1, 30), c(20, 42), c("name", "ssn")))
```

- `fwf_cols()` - Use user provided named widths.
```{r}
vroom_fwf(fwf_sample, fwf_cols(name = 20, state = 10, ssn = 12))
```

- `fwf_cols()` - Use user provided named pairs of positions.
```{r}
vroom_fwf(fwf_sample, fwf_cols(name = c(1, 20), ssn = c(30, 42)))
```

## Column types

vroom guesses the data types of columns as they are read, however sometimes it
is necessary to change the type of one or more columns.

The available specifications are: (with single letter abbreviations in quotes)

* `col_logical()` 'l', containing only `T`, `F`, `TRUE`, `FALSE`, `1` or `0`.
* `col_integer()` 'i', integer values.
* `col_double()` 'd', floating point values.
* `col_number()` [n], numbers containing the `grouping_mark`
* `col_date(format = "")` [D]: with the locale's `date_format`.
* `col_time(format = "")` [t]: with the locale's `time_format`.
* `col_datetime(format = "")` [T]: ISO8601 date times.
* `col_factor(levels, ordered)` 'f', a fixed set of values.
* `col_character()` 'c', everything else.
* `col_skip()` '_, -', don't import this column.
* `col_guess()` '?', parse using the "best" type based on the input.

You can tell vroom what columns to use with the `col_types()` argument in a number of ways.

If you only need to override a single column the most concise way is to use a named vector.

```{r}
# read the 'hp' columns as an integer
vroom(vroom_example("mtcars.csv"), col_types = c(hp = "i"))

# also skip reading the 'cyl' column
vroom(vroom_example("mtcars.csv"), col_types = c(hp = "i", cyl = "_"))

# also read the gears as a factor
vroom(vroom_example("mtcars.csv"), col_types = c(hp = "i", cyl = "_", gear = "f"))
```

However you can also use the `col_*()` functions in a list.

```{r}
vroom(
  vroom_example("mtcars.csv"),
  col_types = list(hp = col_integer(), cyl = col_skip(), gear = col_factor())
)
```

This is most useful when a column type needs additional information, such as
for categorical data when you know all of the levels of a factor.

```{r}
vroom(
  vroom_example("mtcars.csv"),
  col_types = list(gear = col_factor(levels = c(gear = c("3", "4", "5"))))
)
```

## Writing delimited files

Use `vroom_write()` to write delimited files, the default delimiter is tab.

```{r}
vroom_write(mtcars, "mtcars.tsv")
```

```{r, include = FALSE}
unlink("mtcars.tsv")
```

### Writing CSV delimited files

Use the `delim = ','` to write CSV files

```{r}
vroom_write(mtcars, "mtcars.csv", delim = ",")
```

```{r, include = FALSE}
unlink("mtcars.csv")
```

### Writing compressed files

For gzip, bzip2 and xz compression they will be automatically compressed if the
filename ends in `gz`, `bz2` or `xz`.

```{r}
vroom_write(mtcars, "mtcars.tsv.gz")

vroom_write(mtcars, "mtcars.tsv.bz2")

vroom_write(mtcars, "mtcars.tsv.xz")
```

```{r, include = FALSE}
unlink(c("mtcars.tsv.gz", "mtcars.tsv.bz2", "mtcars.tsv.xz"))
```

It is also possible to use other compressors, such as
[pigz](https://zlib.net/pigz/) a parallel gzip implementation,
[lbzip2](http://lbzip2.org/) a parallel bzip2 implementation or
[pixz](https://github.com/vasi/pixz) a parallel xz implementation by using
`pipe()` to create a pipe connection. The parallel versions can be considerably
faster for large output files.

```{r, eval = nzchar(Sys.which("pigz"))}
vroom_write(mtcars, pipe("pigz > mtcars.tsv.gz"))
```

```{r, include = FALSE}
unlink("mtcars.tsv.gz")
```

## Further reading

`vignette("benchmarks")` discusses the performance of vroom, how it compares to
alternatives and how it achieves its results.
